### 2019-06-08 2-Chapters/5-Chapter/chapter5.tex
Done, I changed := to \overset{.}{=}
Lots of small typos.

As reminded above, the model assumes that the number of players $M$ remains fixed during all the bandit game.
However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
The existing previous work on multi-player bandit models are all motivated by possible applications to wireless networks, but most of them assume $M$ to be fixed.
The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the sensing information, as well as in the players' decisions (\ie, collisions):
\begin{equation}\label{eq:5:defRegretDynamicSetting}
    R_T^{\text{dyn}} := \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
\end{equation}

@@

On the other hand, another possibility is to extend ideas developed for the adversarial setting.
In \cite{AlaturLevyKrause19}, the authors essentially consider the piece-wise stationary as an easier case of an adversarial problem.
The recent work \cite{bande2019adversarial} also proposes a decentralized algorithm that can achieve sub-linear regret ($\cO(T^{3/4})$) under an adversarial multi-players bandit model.
\\
\indent
In \cite{AlaturLevyKrause19}, the authors build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf. Algorithm 1), and they apparently discovered the ``communication trick'' independently from \cite{BoursierPerchet18} to use (virtual) communications between players to set up a ``coordinator'' player and $M-1$ ``followers'' (running respectively their Algorithms 2 and 3).
%
Their key algorithmic technique is to imitate the idealized case where there is full communication between the players. Then, to address the no-communication constraint, we enforce the players to keep the same decisions (arms) within long periods of time (blocks). This gives them the chance to coordinate between themselves via a simple protocol that uses collisions as a primitive, yet effective manner of communication.\\
%
\indent
Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as shown in Theorem~4.1
(they used $K$ for $M$ the number of players, and $N$ for $K$ the number of arms).
This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for Exp3 for the single-player adversarial setting in \cite{Auer02NonStochastic}.
Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB, as its regret scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}.\\
\indent
Moreover, at first sight, this idea of ``meta'' arms implies an exponential blow up in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
the authors provide in Section~4.1 and Lemma~4.1 an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} for a good review).
%
In this second work also, we can note the poor experimental section, as the proposed algorithm is only compared against the \MusicalChair{} algorithm, in ``easy'' problems (small number of break-points $\Upsilon_T$).
Despite being more complicated than other approaches, their algorithm ``\texttt{C \& P}'' should not be too hard to implement by ourselves, and studying its empirical performance is an interesting future work.

>>>>

As reminded above, the model assumes that the number of players $M$ remains fixed during all the bandit game.
However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
The existing previous work on multi-player bandit models are all motivated by possible applications to wireless networks, but most of them assume $M$ to be fixed.
The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the sensing information, as well as in the players' decisions (\ie, collisions):
\begin{equation}\label{eq:5:defRegretDynamicSetting}
    R_T^{\text{dyn}} \eqdef \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
\end{equation}

@@

On the other hand, another possibility is to extend ideas developed for the adversarial setting.
In \cite{AlaturLevyKrause19}, the authors essentially consider the piece-wise stationary as an easier case of an adversarial problem.
The recent work \cite{bande2019adversarial} also proposes a decentralized algorithm that can achieve sub-linear regret ($\cO(T^{3/4})$) under an adversarial multi-players bandit model.
\\
\indent
In \cite{AlaturLevyKrause19}, the authors build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf. Algorithm 1), and they apparently discovered the ``communication trick'' independently from \cite{BoursierPerchet18} to use (virtual) communications between players to set up a ``coordinator'' player and $M-1$ ``followers'' (running respectively their Algorithms 2 and 3).
%
Their key algorithmic technique is to imitate the idealized case where there is full communication between the players. Then, to address the no-communication constraint, we enforce the players to keep the same decisions (arms) within long periods of time (blocks). This gives them the chance to coordinate between themselves via a simple protocol that uses collisions as a primitive, yet effective manner of communication.\\
%
\indent
Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as shown in Theorem~4.1
(they used $K$ for $M$ the number of players, and $N$ for $K$ the number of arms).
This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for Exp3 for the single-player adversarial setting in \cite{Auer02NonStochastic}.
Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (while scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).\\
\indent
Moreover, at first sight, this idea of ``meta'' arms implies an exponential blow up in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
the authors provide in Section~4.1 and Lemma~4.1 an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} for a good review).
%
In this second work also, we can note the poor experimental section, as the proposed algorithm is only compared against the \MusicalChair{} algorithm, in ``easy'' problems (small number of break-points $\Upsilon_T$).
Despite being more complicated than other approaches, their algorithm ``\texttt{C \& P}'' should not be too hard to implement by ourselves, and studying its empirical performance is an interesting future work.

### 2019-06-08 5-Nomenclature/Nomenclature.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment
Lots of small typos.

% --------------------------------
% matches Greek Symbols        > G
\nomenclature[G]{$\nu_k$, $\nu_k(t)$, $\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$ (in Chapter~\ref{chapter:6}), arm $k$ for user $j$ (in Chapter~\ref{chapter:5}), arm $i$ in the $k$-th stationary interval (in Chapter~\ref{chapter:6})}
\nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
\nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda},$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
\nomenclature[G]{$\Delta$}{Usually denotes the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}. In Chapter~\ref{chapter:4} in Algorithm~\ref{algo:43:UCBwithDelay} it denotes a delay. In Chapter~\ref{chapter:6}, it can denote different gaps ($\Delta^{\text{opt}}$ and $\Delta^{\text{change}}$)}
\nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\delta$}{Usually denotes a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}. Also denote a confidence level of an algorithm in Section~\ref{sec:6:ChangePointDetector}}
\nomenclature[G]{$\Upsilon_T$}{Number of break-points in a piece-wise stationary MAB problem in Chapter~\ref{chapter:6}. $\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\tau$}{Length of a sliding-window, \eg, in SW-\UCB{} in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\gamma$}{Usually denotes a discount factor, \eg, in D-\UCB{} or D-TS in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, the parameter for the $\varepsilon$-greedy algorithm, or a drift correction parameter for the \CUSUM{} algorithm in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\beta(n,\delta)$}{Usually denotes a threshold for the statistical (GLR) tests in Chapter~\ref{chapter:6}}

@@

% --------------------------------
% matches Roman Symbols        > L
\nomenclature[L]{$t$, $s$, $n$, $r$}{Time step, $t\in[T]$. Chapter~\ref{chapter:6} also uses $s$, $n$ and $r$, \eg, in the $\sup$ of the stopping times or some technical lemmas}
\nomenclature[L]{$T$}{Time horizon, the duration of the bandit game (always $T\geq1$)}
\nomenclature[L]{$T_0$, $T_1$}{Fixed durations of some algorithms based on different phases, \eg, the Musical Chair algorithm from \cite{Rosenski16}}
\nomenclature[L]{$\cA$}{An algorithm, also referred to as a policy or a strategy. $\cA_1,\dots,\cA_N$ denote the $N$ aggregated algorithms in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} and $\cA_1,\dots,\cA_M$ denotes the algorithms of the $M$ players in Chapter~\ref{chapter:5}}
\nomenclature[L]{$A(t)$, $A^j(t)$}{Decision of algorithm $\cA$ at time $t\in[T]$, $A(t)\in[K]$ (from algorithm $\cA$), decision for user $j\in[M]$ in Chapter~\ref{chapter:5} (from algorithm $\cA^j$)}
\nomenclature[L]{$Y_{k,t}$}{Random sample from the arm $k$ at time $t$}
\nomenclature[L]{$\cO_t$}{Vector of observations until time $t$ in Chapter~\ref{chapter:5}}
\nomenclature[L]{$r(t)$, $r^j(t)$}{Reward obtained at time $t$, for user $j$ at time $t$}
\nomenclature[L]{$K$}{Number of arms for multi-armed bandit games}
% \nomenclature[L]{$[T}$, $[K]$, etc}{The set of integers from $1$ to $T$, $[T]=\{1,\dots,T\} = \{ n \in\N: 1 \leq n \leq T\}$}
\nomenclature[L]{$k$}{Usually denotes the $k$-th arm, $k\in[K]$, mainly used in subscripts}
\nomenclature[L]{$M$}{In Chapter~\ref{chapter:4}, denote the maximum number of retransmission of a packet in the ALOHA protocol, and in Chapter~\ref{chapter:5}, denote the number of player for multi-players bandit games}
\nomenclature[L]{$m$}{In Chapter~\ref{chapter:4}, denote the maximum length of the back-off interval after a collision}
\nomenclature[L]{$i$, $j$}{Usually denotes the $i$-th or $j$-th player, $i,j\in[M]$ in Chapter~\ref{chapter:5}}
\nomenclature[L]{$N$}{Usually denotes the number of independent repetitions of the same numerical experiments (\eg, $N=1000$). In Chapter~\ref{chapter:4}, denote the number of IoT (dynamic) devices in the studied network}

@@

\nomenclature[L]{$N_k(t)$ or $T_k(t)$}{Number of samples obtained for arm $k$ at time $t$ ($N_k(t)$ in Chapter~\ref{chapter:4}, $T_k(t)$ in Chapter~\ref{chapter:5})}
\nomenclature[L]{$\widehat{\mu}_k(t)$}{Empirical mean of rewards obtained for arm $k$ at time $t$}
\nomenclature[L]{$I_k(t)$}{Index of arm $k$ at time $t$ for an index policy}
\nomenclature[L]{$\UCB_k(t)$}{Upper-Confidence Bound (\UCB) of arm $k$ at time $t$ for an index policy}
\nomenclature[L]{$g_k^j(t)$, $g^j(t)$}{Index of (arm $k$) at time $t$ of user $j$ for an index policy in Chapter~\ref{chapter:5}}

\nomenclature[L]{$p$}{In Chapter~\ref{chapter:4}, probability of transmission for the devices following a Bernoulli random emission pattern (\eg, $p=10^{-5}$)}
\nomenclature[L]{$S$, $S_k$}{In Chapter~\ref{chapter:4}, total number of \emph{static} devices in the network, resp. in channel $k$}
\nomenclature[L]{$D$, $D_k$}{In Chapter~\ref{chapter:4}, total number of \emph{dynamic} devices in the network, resp. in channel $k$}

\nomenclature[L]{$C_k(t)$, $C^j(t)$}{In Chapters~\ref{chapter:4} and \ref{chapter:5}, collision indicator, on arm $k\in[K]$ or for user $j\in[M]$, at time $j$}

@@

% --------------------------------
% matches Subscripts           > I
\nomenclature[I]{$x_k$}{Usually denotes a variable depending on an arm, for $k\in[K]$}
\nomenclature[I]{$Y_{k,t}$}{Usually denotes a variable depending on an arm, for $k\in[K]$, and on time, for $t\in[T]$}

>>>>

% --------------------------------
% matches Greek Symbols        > G
\nomenclature[G]{$\nu_k$, $\nu_k(t)$, $\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$ (in Chapter~\ref{chapter:6}), arm $k$ for user $j$ (in Chapter~\ref{chapter:5}), arm $i$ in the $k$-th stationary interval (in Chapter~\ref{chapter:6})}
\nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
\nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda},$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
\nomenclature[G]{$\Delta$}{Usually denotes the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}. In Chapter~\ref{chapter:4} in Algorithm~\ref{algo:43:UCBwithDelay} it denotes a delay. In Chapter~\ref{chapter:6}, it can denote different gaps ($\Delta^{\text{opt}}$ and $\Delta^{\text{change}}$)}
\nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\delta$}{Usually denotes a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}. Also denote a confidence level of an algorithm in Section~\ref{sec:6:ChangePointDetector}}
\nomenclature[G]{$\Upsilon_T$}{Number of break-points in a piece-wise stationary MAB problem in Chapter~\ref{chapter:6}. $\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\tau$}{Length of a sliding-window, \eg, in SW-\UCB{} in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\gamma$}{Usually denotes a discount factor, \eg, in D-\UCB{} or D-TS in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, the parameter for the $\varepsilon$-greedy algorithm, or the drift correction parameter for \CUSUM{} in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\beta(n,\delta)$}{Usually denotes a threshold for the statistical (GLR) tests in Chapter~\ref{chapter:6}}

@@

% --------------------------------
% matches Roman Symbols        > L
\nomenclature[L]{$t$, $s$, $n$, $r$}{Time step, $t\in[T]$. Chapter~\ref{chapter:6} also uses $s$, $n$ and $r$, \eg, in the $\sup$ of the stopping times or some technical lemmas}
\nomenclature[L]{$T$}{Time horizon, the duration of the bandit game (always $T\geq1$)}
\nomenclature[L]{$T_0$, $T_1$}{Fixed durations of some algorithms based on different phases, \eg, the Musical Chair algorithm from \cite{Rosenski16}}
\nomenclature[L]{$\cA$}{An algorithm, also referred to as a policy or a strategy. $\cA_1,\dots,\cA_N$ denote the $N$ aggregated algorithms in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} and $\cA_1,\dots,\cA_M$ denotes the algorithms of the $M$ players in Chapter~\ref{chapter:5}}
\nomenclature[L]{$A(t)$, $A^j(t)$}{Decision of algorithm $\cA$ at time $t\in[T]$, $A(t)\in[K]$ (from algorithm $\cA$), decision for user $j\in[M]$ in Chapter~\ref{chapter:5} (from algorithm $\cA^j$)}
\nomenclature[L]{$Y_{k,t}$}{Random sample from the arm $k$ at time $t$}
\nomenclature[L]{$O_t$}{Vector of observations until time $t$ in Chapter~\ref{chapter:5}}
\nomenclature[L]{$r(t)$, $r^j(t)$}{Reward obtained at time $t$, for user $j$ at time $t$}
\nomenclature[L]{$K$}{Number of arms for multi-armed bandit games}
% \nomenclature[L]{$[T}$, $[K]$, etc}{The set of integers from $1$ to $T$, $[T]=\{1,\dots,T\} = \{ n \in\N: 1 \leq n \leq T\}$}
\nomenclature[L]{$k$}{Usually denotes the $k$-th arm, $k\in[K]$, mainly used in subscripts}
\nomenclature[L]{$M$}{Maximum number of retransmission of a packet in the ALOHA protocol in Chapter~\ref{chapter:4}, and number of player for multi-players bandit games in Chapter~\ref{chapter:5}}
\nomenclature[L]{$m$}{Maximum length of the back-off interval after a collision, in Chapter~\ref{chapter:4}}
\nomenclature[L]{$i$, $j$}{Usually denotes the $i$-th or $j$-th player, $i,j\in[M]$ in Chapter~\ref{chapter:5}}
\nomenclature[L]{$N$}{Usually denotes the number of independent repetitions of the same numerical experiments (\eg, $N=1000$). In Chapter~\ref{chapter:4}, denote the number of IoT (dynamic) devices in the studied network}

@@

\nomenclature[L]{$N_k(t)$}{Number of samples obtained for arm $k$ at time $t$}
\nomenclature[L]{$\widehat{\mu}_k(t)$}{Empirical mean of rewards obtained for arm $k$ at time $t$}
\nomenclature[L]{$I_k(t)$}{Index of arm $k$ at time $t$ for an index policy}
\nomenclature[L]{$\UCB_k(t)$}{Upper-Confidence Bound (\UCB) of arm $k$ at time $t$ for an index policy}
\nomenclature[L]{$g_k^j(t)$, $g^j(t)$}{Index of (arm $k$) at time $t$ of user $j$ for an index policy in Chapter~\ref{chapter:5}}

\nomenclature[L]{$p$}{In Chapter~\ref{chapter:4}, probability of transmission for the devices following a Bernoulli random emission pattern (\eg, $p=10^{-5}$)}
\nomenclature[L]{$S$, $S_k$}{In Chapter~\ref{chapter:4}, total number of \emph{static} devices in the network or in channel $k$}
\nomenclature[L]{$D$, $D_k$}{In Chapter~\ref{chapter:4}, total number of \emph{dynamic} devices in the network or in channel $k$}

\nomenclature[L]{$C_k(t)$, $C^j(t)$}{In Chapter~\ref{chapter:5}, collision indicator on arm $k\in[K]$ or for user $j\in[M]$, at time $t$}

@@

% --------------------------------
% matches Subscripts           > I
\nomenclature[I]{$x_k$}{Usually denotes a variable depending on an arm, for $k\in[K]$}
\nomenclature[I]{$Y_{k,t}$}{Usually denotes a variable depending on an arm $k\in[K]$ and on time $t\in[T]$}
